---
title: "tna v1.1.0 — Debug & Test Report"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 6
)
library(tna)
```

# Package Info

```{r}
cat("tna version:", as.character(packageVersion("tna")), "\n")
cat("R version:", R.version.string, "\n")
```

# Building Models

## build_model (explicit type)

```{r}
model <- build_model(group_regulation, type = "relative")
print(model)
```

## tna() shortcut (type = "relative")

```{r}
model <- tna(group_regulation)
summary(model)
```

## ftna() shortcut (type = "frequency")

```{r}
model_f <- ftna(group_regulation)
print(model_f)
```

## ctna() shortcut (type = "co-occurrence")

```{r}
model_c <- ctna(group_regulation)
print(model_c)
```

## atna() shortcut (type = "attention")

```{r}
model_a <- atna(group_regulation)
print(model_a)
```

## group_model from mhmm

```{r}
gmodel <- group_model(engagement_mmm)
print(gmodel)
```

## group_model with manual groups

```{r}
group <- c(rep("High", 1000), rep("Low", 1000))
gmodel_manual <- group_model(group_regulation, group = group)
summary(gmodel_manual)
```

## sna (edge list)

```{r}
set.seed(42)
d <- data.frame(
  from = sample(LETTERS[1:4], 100, replace = TRUE),
  to = sample(LETTERS[1:4], 100, replace = TRUE),
  weight = rexp(100)
)
model_sna <- sna(d)
```

# Plotting

## plot.tna

```{r}
model <- tna(group_regulation)
plot(model)
```

## plot with spring layout and scaled nodes

```{r}
plot(model, layout = "spring", scale_nodes = "OutStrength")
```

## plot group_tna

```{r}
gmodel <- group_model(engagement_mmm)
plot(gmodel)
```

## plot_frequencies

```{r}
model <- tna(group_regulation)
plot_frequencies(model)
```

## plot_frequencies (group_tna)

```{r}
gmodel <- group_model(engagement_mmm)
plot_frequencies(gmodel)
```

## plot_mosaic

```{r}
model_f <- ftna(group_regulation)
plot_mosaic(model_f)
```

## plot_mosaic (group_tna)

```{r}
gmodel <- group_model(engagement_mmm, type = "frequency")
plot_mosaic(gmodel)
```

## plot_sequences (index)

```{r}
plot_sequences(group_regulation)
```

## plot_sequences (distribution)

```{r}
plot_sequences(group_regulation, type = "distribution")
```

## plot_sequences (group_tna)

```{r}
gmodel <- group_model(engagement_mmm)
plot_sequences(gmodel)
```

## plot_compare

```{r}
model_a <- tna(group_regulation[1:1000, ])
model_b <- tna(group_regulation[1001:2000, ])
plot_compare(model_a, model_b)
```

## plot_associations

```{r}
model_f <- ftna(group_regulation)
plot_associations(model_f)
```

## hist

```{r}
model <- tna(group_regulation)
hist(model)
```

# Data Preparation

## prepare_data

```{r}
results <- prepare_data(
  group_regulation_long,
  actor = "Actor", time = "Time", action = "Action"
)
print(results$sequence_data)
print(results$statistics)
```

## simulate

```{r}
model <- tna(group_regulation)
sim <- simulate(model, nsim = 5, seed = 123, max_len = 10)
print(sim)
```

# Centrality Measures

## centralities (tna)

```{r}
model <- tna(group_regulation)
cm <- centralities(model)
print(cm)
```

## centralities (group_tna)

```{r}
gmodel <- group_model(engagement_mmm)
cm_g <- centralities(gmodel)
print(cm_g)
```

## plot centralities

```{r}
model <- tna(group_regulation)
cm <- centralities(model)
plot(cm, ncol = 3, reorder = TRUE)
```

## specific measures

```{r}
centralities(model, measures = c("OutStrength", "Betweenness"))
```

## normalized

```{r}
centralities(model, normalize = TRUE)
```

## betweenness_network

```{r}
bn <- betweenness_network(model)
print(bn)
```

## estimate_cs

```{r}
cs <- estimate_cs(model,
  measures = c("InStrength", "OutStrength"),
  iter = 100
)
print(cs)
plot(cs)
```

# Network Structure

## communities (tna)

```{r}
model <- tna(group_regulation)
comm <- communities(model)
print(comm)
```

## plot communities

```{r}
plot(comm, method = "spinglass")
```

## communities (group_tna)

```{r}
gmodel <- group_model(engagement_mmm)
comm_g <- communities(gmodel)
print(comm_g)
```

## cliques

```{r}
model <- tna(group_regulation)
cliq <- cliques(model, size = 2)
print(cliq)
```

## plot cliques

```{r}
plot(cliq, n = 3, ask = FALSE)
```

# Comparison

## compare (group_tna)

```{r}
gmodel <- group_model(engagement_mmm)
comp <- compare(gmodel, i = 1, j = 2)
print(comp)
```

## plot comparison

```{r}
plot(comp)
```

## compare_sequences (group_tna)

```{r}
gmodel <- group_model(engagement_mmm)
comp_seq <- compare_sequences(gmodel)
print(head(comp_seq, 10))
```

## plot compare_sequences

```{r}
plot(comp_seq)
```

## compare_sequences (manual groups)

```{r}
group <- c(rep("High", 1000), rep("Low", 1000))
comp_manual <- compare_sequences(group_regulation, group)
print(head(comp_manual, 10))
```

## permutation_test

```{r}
model_x <- tna(group_regulation[1:200, ])
model_y <- tna(group_regulation[1001:1200, ])
perm <- permutation_test(model_x, model_y, iter = 100)
print(perm)
```

## plot permutation_test

```{r}
plot(perm)
```

# Validation

## bootstrap

```{r}
model <- tna(group_regulation)
boot <- bootstrap(model, iter = 100)
print(boot)
```

## summary bootstrap

```{r}
summary(boot)
```

## plot bootstrap

```{r}
plot(boot)
```

## bootstrap_cliques

```{r}
bc <- bootstrap_cliques(model, size = 2, iter = 100)
print(bc)
```

## prune (threshold)

```{r}
model <- tna(group_regulation)
pruned_t <- prune(model, method = "threshold", threshold = 0.1)
print(pruned_t)
```

## prune (lowest)

```{r}
pruned_p <- prune(model, method = "lowest", lowest = 0.05)
print(pruned_p)
```

## prune (disparity)

```{r}
pruned_d <- prune(model, method = "disparity", level = 0.5)
print(pruned_d)
```

## pruning_details

```{r}
pruning_details(pruned_t)
```

## deprune + reprune

```{r}
restored <- deprune(pruned_t)
cat("Deprune OK\n")
repruned <- reprune(restored)
cat("Reprune OK\n")
```

## plot pruned network

```{r}
plot(pruned_t)
```

# Clustering

## cluster_sequences

```{r}
result <- cluster_sequences(group_regulation[1:200, ], k = 3, dissimilarity = "osa")
print(result)
```

## group_model from clustering

```{r}
gmodel_clust <- group_model(result)
print(gmodel_clust)
plot(gmodel_clust)
```

## rename_groups

```{r}
gmodel <- group_model(engagement_mmm)
gmodel_renamed <- rename_groups(gmodel, c("A", "B", "C"))
cat("Original:", names(group_model(engagement_mmm)), "\n")
cat("Renamed:", names(gmodel_renamed), "\n")
```

## mmm_stats

```{r}
mmm_stats(engagement_mmm)
```

# Summary & Conversion

## summary (tna)

```{r}
model <- tna(group_regulation)
summary(model)
```

## summary (group_tna)

```{r}
gmodel <- group_model(engagement_mmm)
summary(gmodel)
```

## as.igraph

```{r}
model <- tna(group_regulation)
g <- as.igraph(model)
print(g)
```
